---
title: "Hw3"
author: "Ashton Pallottini"
date: "10/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Number 1

####Number 1.1

```{r}
answer1.1 <- as.data.frame(matrix(nrow = 4, ncol = 3))
colnames(answer1.1) <- c("Gamete", "Expected Frequency (LE)", "Observed Frequency")
answer1.1[,1] <- c("AB", "Ab", "aB", "ab")
answer1.1[,2] <- c(.2*.3, .2*.7, .8*.3, .8*.7)
answer1.1[,3] <- c(.1, .2-.1, .3-.1, .8-(.3-.1))
answer1.1
```

####Number 1.2

```{r}
#D = P(AB) - P(A)P(B)
D <- answer1.1$`Observed Frequency`[1] - (.2*.3)
D

#D' = |D|/min(P(A)P(b), P(B)P(a))
D.prime <- abs(D)/min(.2*.7, .8*.3)
D.prime

#R^2 = D^2/(P(A)*P(b)*P(a)*P(B))
r.sq <- D^2/(.2*.7*.8*.3)
r.sq
```

####Number 1.3

```{r}
chi.sq <- sum((100*answer1.1[,3] - 100*answer1.1[,2])^2/(100*answer1.1[,2]))
chi.sq

pchisq(chi.sq, df = 1, lower.tail = F)
```

We have Ho: A and B are in LE versus Ha: A and B are not in LE. We get a chi-squared statistic of 4.762. The degree of freedom is one, which yields a p value of 0.0291. We have enough evidence to say that this is in linkage disequilibrium, and we reject Ho at the 5% significance level.

####Number 1.4

```{r}
answer1.4 <- as.data.frame(matrix(nrow = 9, ncol = 2))
colnames(answer1.4) <- c("Genotype", "Expected Frequency")
answer1.4[,1] <- c("AA BB", "AA Bb", "AA bb", "Aa BB", "Aa Bb", "Aa bb", "aa BB", "aa Bb", "aa bb")
answer1.4[,2] <- c(.06^2, 2*.06*.14, .14^2, 2*.06*.24, 2*(.06*.56+.14*.24), 2*.14*.56, .24^2, 2*.24*.56, .56^2)
answer1.4
```

###Number 2

####Number 2.1

```{r}
2*(.5)^(6-1)
```

There are two loops here, each with Q = 6. Thus, the coancestry is 2*(1/2)^(6-1) = (1/2)^4 = 1/16 = 0.0625.

####Number 2.2

```{r}
2*(.5)^(6-1)
```

There are two loops here, each with Q = 6. Thus, the coancestry is 2*(1/2)^(6-1) = (1/2)^4 = 1/16 = 0.0625. Also, the common ancestor between the two is B. If we take the inbreeding of B to equal zero, then the inbreeding for both pups will equal the coancestry. As such, Inbreeding Pup 1 = Inbreeding Pup 2 = Coancestry = 0.0625.

###Number 3

####Number 3.1

```{r}
var.x1i.a1 <- 1^2*(2*.2*(1-.2))
var.x1i.a1

var.x2i.a2 <- 2^2*(2*.3*(1-.3))
var.x2i.a2
```

For both of the above situations, var(ax) = (a^2)var(x) = (a^2)(2p(1-p)), which yields variances of 0.32 and 1.68, respectively.

####Number 3.2

```{r}
this.cov <- var.x1i.a1 + var.x2i.a2 + 2*1*2*.2
this.cov
```

We have var(x1i.a1 + x2i.a2) = var(x1i.a1) + var(x2i.a2) + 2cov(x1i.a1, x2i.a2) = 0.32 + 1.68 + 2a1a2(cov(x1i, x2i)) which equals 2.8 when we substitute everything in.

####Number 4

```{r}
var.x1i.a1 <- 1^2*(2*.2*(1-.2))
var.x1i.a1

var.x2i.a2 <- 2^2*(2*.2*(1-.2))
var.x2i.a2
```

For both of the above situations, var(ax) = (a^2)var(x) = (a^2)(2p(1-p)), which yields variances of 0.32 and 1.28, respectively. Also, cov(u1i, u2i) = a1a2(cov(x1i, x2i)) = 2(cov(x1i, x2i)).
